---
title: "Madison Lake"
author: "Felicia Cruz"
date: "8/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(ggbeeswarm)
library(lubridate)
```

```{r}
ice <- read_csv(here("ntl33_v7.csv"), na = "-999") %>% 
  clean_names()

met_data <- read_csv(here("ntl20_v6.csv"), na = "-999")

```

```{r}
ggplot(ice, aes(x = lakeid, y = ice_duration)) +
  geom_jitter(aes(color = lakeid), show.legend = FALSE) +
  geom_boxplot(aes(color = lakeid),
               alpha = 0.5,
               outlier.shape = NA,
               show.legend = FALSE)

```

```{r}
mean_ice_dur_year <- ice %>% 
  group_by(year4) %>% 
  summarize(mean_ice_dur = mean(ice_duration, na.rm = TRUE))

```

```{r}
ggplot(mean_ice_dur_year, aes(x = year4, y = mean_ice_dur)) +
  geom_line()
```

In the line graph above, we can see an overall decrease in average ice duration (days) from 1850-2019. 


```{r}
air_temp_avg <- met_data %>% 
  filter(month %in% c(12, 1, 2)) %>% 
  group_by(year4) %>% 
  summarize(mean_air_temp = mean(ave_air_temp_adjusted))


ggplot(air_temp_avg, aes(x = year4, y = mean_air_temp)) +
  geom_line()
```

The line graph above shows a general increase in average daily air temperature (degrees Celsius) from 1869-2012. 

```{r}
temp_and_ice <- left_join(air_temp_avg, mean_ice_dur_year, by = "year4")


ggplot(temp_and_ice, aes(x = mean_air_temp, y = mean_ice_dur)) +
  geom_point()
```

This scatterplot shows a negative correlation between mean air temperature and mean ice duration. As the average daily air temperature increases, the average ice duration generally decreases. 

```{r}
subset_ice <- ice %>% 
  filter(year4 >= 1970) %>% 
  mutate(ice_off = lubridate::ymd(ice_off)) %>% 
  mutate(day_off = lubridate::yday(ice_off))

ggplot(subset_ice, aes(x = year4, y = day_off)) +
  geom_line() +
  geom_point() +
  labs(title = "Ice Thaw Day",
       subtitle = "Madison Lake Area 1869 - 2019",
       x = "Year", 
       y = "Day") +
  theme_minimal()

```

There is no immediately obvious trend between year and ice thaw day. There is a lot of fluxuation over the years.

